---
title: "R Notebook"
output: 
  html_document:
    keep_md: yes
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme: cerulean #flatly #journal #cosmo
  pdf_document: default
---


# Intro

This markdown describes the data cleaning/wrangling steps to process and visualise the bed data from 2020 onwards. The 99.9% onfidence intervals for the mean value were calculated using the bootstrap distribution of the bed data from 2016 to 2019 (refer to markdown document ``my_phs_testpad_aux.Rmd``).


# Preamble

## Call libraries
```{r}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggthemes)
```


## Set up colour palette
```{r}
# The palette as per SF:
phs_palette <- c("#99DAF5", "#004785", "#C027B9", "#82BB25")

# # To use for fills, add
#   scale_fill_manual(values = phs_palette)
# 
# # To use for line and point colors, add
#   scale_colour_manual(values = phs_palette)
```


# Number of beds - Data Wrangling

## View bed data
```{r}
beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv") 


glimpse(beds)
```



### Check data  for NA values
```{r}
quarters_in_scope <- c("2020Q1", "2020Q2", "2020Q3", "2020Q4", "2021Q1", "2021Q2", "2021Q3")

beds %>%
  filter(Quarter %in% quarters_in_scope) %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

### Keep data related to our time frame of interest 
```{r}
quarters_in_scope <- c("2020Q1", "2020Q2", "2020Q3", "2020Q4", "2021Q1", "2021Q2", "2021Q3")
```


```{r}
beds <- beds %>%
  filter(Quarter %in% quarters_in_scope)
```



```{r}
glimpse(beds)
```



### How many HB items do we have?
```{r}
beds %>%
  distinct(HB)
```


### Keep SpecialtyName == All Acute for the analysis
```{r}
beds <- beds %>%
  filter(SpecialtyName == "All Acute")
```


## View health board data
```{r}
health_board_labels <- read_csv("data/hb14_hb19.csv") 

glimpse(health_board_labels)
```

### Remove junk
```{r}
health_board_labels <- health_board_labels %>% 
  select(HB, HBName)
  

glimpse(health_board_labels)
```



## View special health board data

**Note:** I had to use the following arguments to read the csv file, because the "The Golden Jubilee National Hospital" in the SHName column had a special Unicode character that I couldn't get rid off (if I kept the unicode character, I couldn't ``glimpse`` the data).
```{r}
shb_labels <- read_csv("data/special-health-boards_19022021.csv", locale = locale(encoding = "ASCII")) 

glimpse(shb_labels)
```
### Remove junk
```{r}
shb_labels <- shb_labels %>% 
  select(SHB, SHBName)
  

glimpse(shb_labels)
```



### View country label data
```{r}
country_labels <- read_csv("data/geography_codes_and_labels_country.csv") 


glimpse(country_labels)
```


# Join tables 

## Join bed & special health board
```{r}
beds <- beds %>%
  left_join(shb_labels, by = c("HB" = "SHB"))

glimpse(beds)
```

```{r}
beds %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



## Join bed & health board
```{r}
beds <- beds %>%
  left_join(health_board_labels, "HB")

glimpse(beds)
```

```{r}
beds %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



```{r}
# beds <- beds %>% 
#   mutate(Country = if_else(!is.na(Country.x), Country.x, Country.y)) %>% 
#   select(-Country.x, -Country.y)
# 
# glimpse(beds) 
```



## Join bed & country labels
```{r}
beds <- beds %>%
  left_join(country_labels, by = c("HB" = "Country"))

glimpse(beds)
```



```{r}
beds %>%
  # filter(Quarter %in% quarters_in_scope) %>%
  # filter(SpecialtyName == "All Acute") %>% 
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

##  Tidy up finalised bed table
```{r}
beds <- beds %>%
  clean_names()

glimpse(beds)
```


**Sanity Check - There should be on average 13,226 available staffed beds in Scotland during quarter ending September 2021 (PHS report)**
```{r}
beds %>%
  filter(country_name == "Scotland")

```



# Add confidence interval information in beds 


## View health board data
```{r}
hb_ci <- read_csv("hb_ci.csv") 

glimpse(hb_ci)
```

## View special health board data
```{r}
shb_ci <- read_csv("shb_ci.csv") 

glimpse(shb_ci)
```

## View country data
```{r}
country_ci <- read_csv("country_ci.csv") 

glimpse(country_ci)
```


## Join bed & country ci
```{r}
country_ci <- country_ci%>% 
  select(-country_name)

beds <- beds %>%
  left_join(country_ci, by = c("hb" = "country"))

glimpse(beds)
```

## Join bed & health board
```{r}
hb_ci <- hb_ci%>% 
  select(-hb_name)


beds <- beds %>%
  left_join(hb_ci, "hb")

glimpse(beds)
```

## Clean up junk
```{r}
beds <- beds %>%
  mutate(lower_tob_ci999 = if_else(!is.na(lower_tob_ci999.x), lower_tob_ci999.x, lower_tob_ci999.y)) %>%
  select(-lower_tob_ci999.x, -lower_tob_ci999.y) %>%
  mutate(upper_tob_ci999 = if_else(!is.na(upper_tob_ci999.x), upper_tob_ci999.x, upper_tob_ci999.y)) %>%
  select(-upper_tob_ci999.x, -upper_tob_ci999.y) %>%
  mutate(lower_aasb_ci999 = if_else(!is.na(lower_aasb_ci999.x), lower_aasb_ci999.x, lower_aasb_ci999.y)) %>%
  select(-lower_aasb_ci999.x, -lower_aasb_ci999.y) %>%
  mutate(upper_aasb_ci999 = if_else(!is.na(upper_aasb_ci999.x), upper_aasb_ci999.x, upper_aasb_ci999.y)) %>%
  select(-upper_aasb_ci999.x, -upper_aasb_ci999.y) %>%
  mutate(lower_aob_ci999 = if_else(!is.na(lower_aob_ci999.x), lower_aob_ci999.x, lower_aob_ci999.y)) %>%
  select(-lower_aob_ci999.x, -lower_aob_ci999.y) %>%
  mutate(upper_aob_ci999 = if_else(!is.na(upper_aob_ci999.x), upper_aob_ci999.x, upper_aob_ci999.y)) %>%
  select(-upper_aob_ci999.x, -upper_aob_ci999.y) %>%
  mutate(lower_po_ci999 = if_else(!is.na(lower_po_ci999.x), lower_po_ci999.x, lower_po_ci999.y)) %>%
  select(-lower_po_ci999.x, -lower_po_ci999.y) %>%
  mutate(upper_po_ci999 = if_else(!is.na(upper_po_ci999.x), upper_po_ci999.x, upper_po_ci999.y)) %>%
  select(-upper_po_ci999.x, -upper_po_ci999.y)

glimpse(beds)
```

## Join bed & special health board
```{r}
shb_ci <- shb_ci%>% 
  select(-shb_name)


beds <- beds %>%
  left_join(shb_ci, , by = c("hb" = "shb"))

glimpse(beds)
```


## Clean up junk
```{r}
beds <- beds %>%
  mutate(lower_tob_ci999 = if_else(!is.na(lower_tob_ci999.x), lower_tob_ci999.x, lower_tob_ci999.y)) %>%
  select(-lower_tob_ci999.x, -lower_tob_ci999.y) %>%
  mutate(upper_tob_ci999 = if_else(!is.na(upper_tob_ci999.x), upper_tob_ci999.x, upper_tob_ci999.y)) %>%
  select(-upper_tob_ci999.x, -upper_tob_ci999.y) %>%
  mutate(lower_aasb_ci999 = if_else(!is.na(lower_aasb_ci999.x), lower_aasb_ci999.x, lower_aasb_ci999.y)) %>%
  select(-lower_aasb_ci999.x, -lower_aasb_ci999.y) %>%
  mutate(upper_aasb_ci999 = if_else(!is.na(upper_aasb_ci999.x), upper_aasb_ci999.x, upper_aasb_ci999.y)) %>%
  select(-upper_aasb_ci999.x, -upper_aasb_ci999.y) %>%
  mutate(lower_aob_ci999 = if_else(!is.na(lower_aob_ci999.x), lower_aob_ci999.x, lower_aob_ci999.y)) %>%
  select(-lower_aob_ci999.x, -lower_aob_ci999.y) %>%
  mutate(upper_aob_ci999 = if_else(!is.na(upper_aob_ci999.x), upper_aob_ci999.x, upper_aob_ci999.y)) %>%
  select(-upper_aob_ci999.x, -upper_aob_ci999.y) %>%
  mutate(lower_po_ci999 = if_else(!is.na(lower_po_ci999.x), lower_po_ci999.x, lower_po_ci999.y)) %>%
  select(-lower_po_ci999.x, -lower_po_ci999.y) %>%
  mutate(upper_po_ci999 = if_else(!is.na(upper_po_ci999.x), upper_po_ci999.x, upper_po_ci999.y)) %>%
  select(-upper_po_ci999.x, -upper_po_ci999.y)

glimpse(beds)
```


# KPI plots (Scotland: 2020Q1 - 2021Q3)


```{r}
## Test plot

beds %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, group = country_name) +
  geom_line(aes(y = total_occupied_beds, colour = country_name)) +
  geom_point(aes(y = total_occupied_beds, colour = country_name)) +
  geom_line(aes(y = lower_tob_ci999), colour = "red") +
  geom_line(aes(y = upper_tob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))
  
```


## Selected Geography: Scotland
```{r}
beds %>%
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, group = country_name) +
  geom_line(aes(y = total_occupied_beds, colour = country_name)) +
  geom_point(aes(y = total_occupied_beds, colour = country_name)) +
  geom_line(aes(y = lower_tob_ci999), colour = "red") +
  geom_line(aes(y = upper_tob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, group = country_name) +
  geom_line(aes(y = average_available_staffed_beds, colour = country_name)) +
  geom_point(aes(y = average_available_staffed_beds, colour = country_name)) +
  geom_line(aes(y = lower_aasb_ci999), colour = "red") +
  geom_line(aes(y = upper_aasb_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, group = country_name) +
  geom_line(aes(y = average_occupied_beds, colour = country_name)) +
  geom_point(aes(y = average_occupied_beds, colour = country_name)) +
  geom_line(aes(y = lower_aob_ci999), colour = "red") +
  geom_line(aes(y = upper_aob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, group = country_name) +
  geom_line(aes(y = percentage_occupancy, colour = country_name)) +
  geom_point(aes(y = percentage_occupancy, colour = country_name)) +
  geom_line(aes(y = lower_po_ci999), colour = "red") +
  geom_line(aes(y = upper_po_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

```

### Sanity check No. 1. 

Check if I pick in the location column the aggregated results of value S08000031 for NHS Greater Glasgow and S08000020 for Grampian
```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(hb_name == "NHS Grampian" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

## Selected Geography: NHS Board
```{r}
beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = lower_tob_ci999), colour = "red") +
  geom_line(aes(y = upper_tob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = lower_aasb_ci999), colour = "red") +
  geom_line(aes(y = upper_aasb_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = lower_aob_ci999), colour = "red") +
  geom_line(aes(y = upper_aob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_point(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_line(aes(y = lower_po_ci999), colour = "red") +
  geom_line(aes(y = upper_po_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

```


### Sanity check No. 2.
Check to see if I pick in the location column the aggregated results of value SB0801 
```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(shb_name == "The Golden Jubilee National Hospital" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

```{r}
beds %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



## Selected Geography: Special Health Board
```{r}
beds %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, group = shb_name) +
  geom_line(aes(y = total_occupied_beds, colour = shb_name)) +
  geom_point(aes(y = total_occupied_beds, colour = shb_name)) +
  geom_line(aes(y = lower_tob_ci999), colour = "red") +
  geom_line(aes(y = upper_tob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%
  ggplot() +
  aes(x = quarter, group = shb_name) +
  geom_line(aes(y = average_available_staffed_beds, colour = shb_name)) +
  geom_point(aes(y = average_available_staffed_beds, colour = shb_name)) +
  geom_line(aes(y = lower_aasb_ci999), colour = "red") +
  geom_line(aes(y = upper_aasb_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%
  ggplot() +
  aes(x = quarter, group = shb_name) +
  geom_line(aes(y = average_occupied_beds, colour = shb_name)) +
  geom_point(aes(y = average_occupied_beds, colour = shb_name)) +
  geom_line(aes(y = lower_aob_ci999), colour = "red") +
  geom_line(aes(y = upper_aob_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, group = shb_name) +
  geom_line(aes(y = percentage_occupancy, colour = shb_name)) +
  geom_point(aes(y = percentage_occupancy, colour = shb_name)) +
  geom_line(aes(y = lower_po_ci999), colour = "red") +
  geom_line(aes(y = upper_po_ci999), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

```


# Save cleaned data

Write updated beds data to a CSV file.
```{r}
 write_csv(beds, here("beds_ci_clean.csv"))
```




# Doodles

```{r}
country_name_beds <- unique(beds$country_name) %>% 
  discard(is.na)
  
```

```{r}
hb_name_beds <- unique(beds$hb_name) %>% 
  discard(is.na)
  
```


```{r}
shb_name_beds <- unique(beds$shb_name) %>% 
  discard(is.na)
  
```





















