---
title: "R Notebook"
output: 
  html_document:
    keep_md: yes
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme: cerulean #flatly #journal #cosmo
  pdf_document: default
---


# Intro

This markdown describes the data cleaning/wrangling steps to process and visualise the bed data from 2020 onwards. The mean value of the bed data from 2016 to 2019 is also calculated in order to provide a reference value for comparisons.


# Preamble

## Call libraries
```{r}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggthemes)
```


## Set up colour palette
```{r}
# The palette as per SF:
phs_palette <- c("#99DAF5", "#004785", "#C027B9", "#82BB25")

# # To use for fills, add
#   scale_fill_manual(values = phs_palette)
# 
# # To use for line and point colors, add
#   scale_colour_manual(values = phs_palette)
```


# Number of beds - Data Wrangling

## View bed data
```{r}
beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv") 


glimpse(beds)
```



### Check data  for NA values
```{r}
quarters_in_scope <- c("2020Q1", "2020Q2", "2020Q3", "2020Q4", "2021Q1", "2021Q2", "2021Q3")

beds %>%
  filter(Quarter %in% quarters_in_scope) %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

### Keep data related to our time frame of interest 
```{r}
quarters_in_scope <- c("2020Q1", "2020Q2", "2020Q3", "2020Q4", "2021Q1", "2021Q2", "2021Q3")
```


```{r}
beds <- beds %>%
  filter(Quarter %in% quarters_in_scope)
```



```{r}
glimpse(beds)
```



### How many HB items do we have?
```{r}
beds %>%
  distinct(HB)
```


### Keep "SpecialtyName == ``All Acute``" for the analysis
```{r}
beds <- beds %>%
  filter(SpecialtyName == "All Acute")
```


## View health board data
```{r}
health_board_labels <- read_csv("data/hb14_hb19.csv") 

glimpse(health_board_labels)
```

### View all HB values
```{r}
health_board_labels %>% 
  distinct(HB, HBName, HBDateEnacted)
```

### Remove junk
```{r}
health_board_labels <- health_board_labels %>% 
  select(HB, HBName)
  

glimpse(health_board_labels)
```



## View special health board data

**Note:** I had to use the following arguments to read the csv file, because the "The Golden Jubilee National Hospital" in the SHName column had a special Unicode character that I couldn't get rid off (if I kept the unicode character, I couldn't ``glimpse`` the data).
```{r}
shb_labels <- read_csv("data/special-health-boards_19022021.csv", locale = locale(encoding = "ASCII")) 

glimpse(shb_labels)
```

### View all SHB values
```{r}
shb_labels %>% 
  distinct(SHB, SHBName)
```

### Remove junk
```{r}
shb_labels <- shb_labels %>% 
  select(SHB, SHBName)
  

glimpse(shb_labels)
```



### View country label data
```{r}
country_labels <- read_csv("data/geography_codes_and_labels_country.csv") 


glimpse(country_labels)
```


# Join tables 

## Join bed & special health board
```{r}
beds <- beds %>%
  left_join(shb_labels, by = c("HB" = "SHB"))

glimpse(beds)
```

```{r}
beds %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



## Join bed & health board
```{r}
beds <- beds %>%
  left_join(health_board_labels, "HB")

glimpse(beds)
```

```{r}
beds %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```


## Join bed & country labels
```{r}
beds <- beds %>%
  left_join(country_labels, by = c("HB" = "Country"))

glimpse(beds)
```


```{r}
beds %>%
  # filter(Quarter %in% quarters_in_scope) %>%
  # filter(SpecialtyName == "All Acute") %>% 
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

## Concatenate ``SHBName``, ``HBName`` and ``CountryName`` in a single column called ``HBName``
```{r}
beds <- beds %>%
  mutate(HBName = if_else(!is.na(HBName), HBName, SHBName)) %>%
  mutate(HBName = if_else(!is.na(HBName), HBName, CountryName)) %>%
  select(-SHBName, -CountryName)

glimpse(beds)
```


##  Tidy up finalised bed table
```{r}
beds <- beds %>%
  clean_names()

glimpse(beds)
```


**Sanity Check - There should be on average 13,226 available staffed beds in Scotland during quarter ending September 2021 (PHS report)**
```{r}
beds %>%
  filter(hb_name == "Scotland")

```

# KPI plots (Scotland: 2020Q1 - 2021Q3)


## Selected Geography: Scotland
```{r}
beds %>%
  filter(hb_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

beds %>%
  filter(hb_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))
```

### Sanity check No. 1. 

Check if I pick in the location column the aggregated results of value S08000031 for NHS Greater Glasgow and S08000020 for Grampian
```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(hb_name == "NHS Grampian" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

## Selected Geography: NHS Board
```{r}
beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

beds %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))
```


### Sanity check No. 2.
Check to see if I pick in the location column the aggregated results of value SB0801 
```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(hb_name == "The Golden Jubilee National Hospital" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

```{r}
beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



## Selected Geography: Special Health Board
```{r}
beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))
```



# KPI reference values

Calculate reference bed values (calculated averaged values) from pre-covid data (2016 to 2019) 

## View bed data
```{r}
beds_oos <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv") 

```

### Check data  for NA values
```{r}
quarters_out_of_scope <- c("2016Q3", "2016Q4", "2017Q1", "2017Q2", "2017Q3",
                           "2017Q4", "2018Q1", "2018Q2", "2018Q3", "2018Q4", 
                           "2019Q1", "2019Q2", "2019Q3", "2019Q4")

beds_oos %>%
  filter(Quarter %in% quarters_out_of_scope) %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```


### Keep data related to our time frame of interest 
```{r}
quarters_out_of_scope <- c("2016Q3", "2016Q4", "2017Q1", "2017Q2", "2017Q3",
                           "2017Q4", "2018Q1", "2018Q2", "2018Q3", "2018Q4", 
                           "2019Q1", "2019Q2", "2019Q3", "2019Q4")
```


```{r}
beds_oos <- beds_oos %>%
  filter(Quarter %in% quarters_out_of_scope)
```



```{r}
glimpse(beds_oos)
```


### How many HB items do we have?
```{r}
beds_oos %>%
  distinct(HB)
```


### Keep SpecialtyName == All Acute for the analysis
```{r}
beds_oos <- beds_oos %>%
  filter(SpecialtyName == "All Acute")
```


## Join tables 

### Join bed & special health board
```{r}
beds_oos <- beds_oos %>%
  left_join(shb_labels, by = c("HB" = "SHB"))

glimpse(beds_oos)
```

```{r}
beds_oos %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



### Join bed & health board
```{r}
beds_oos <- beds_oos %>%
  left_join(health_board_labels, "HB")

glimpse(beds_oos)
```

```{r}
beds_oos %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



### Join bed & country labels
```{r}
beds_oos <- beds_oos %>%
  left_join(country_labels, by = c("HB" = "Country"))

glimpse(beds_oos)
```



```{r}
beds_oos %>%
  # filter(Quarter %in% quarters_in_scope) %>%
  # filter(SpecialtyName == "All Acute") %>% 
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```


### Concatenate ``SHBName``, ``HBName`` and ``CountryName`` in a single column called ``HBName``
```{r}
beds_oos <- beds_oos %>%
  mutate(HBName = if_else(!is.na(HBName), HBName, SHBName)) %>%
  mutate(HBName = if_else(!is.na(HBName), HBName, CountryName)) %>%
  select(-SHBName, -CountryName)

glimpse(beds_oos)
```


###  Tidy up finalised bed table
```{r}
beds_oos <- beds_oos %>%
  clean_names()

glimpse(beds_oos)
```

**Sanity Check - There should be on average 13,270 available staffed beds in Scotland during quarter ending December 2019 (PHS report)**
```{r}
beds_oos %>%
  # filter(hb_name == "NHS Borders")
  filter(hb_name == "Scotland")

```

## KPI plots (Scotland: 2016Q3 - 2019Q4)


### Selected Geography: Scotland
```{r}
beds_oos %>%
  filter(hb_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))

beds_oos %>%
  filter(hb_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))
```


### Sanity check No. 1. 

Check if I pick the aggregated results of value S08000031 for NHS Greater Glasgow and S08000020 for Grampian in the location column
```{r}
beds_oos %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



```{r}
beds_oos %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(hb_name == "NHS Grampian" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```


### Selected Geography: NHS Board
```{r}
beds_oos %>%
  filter(hb_name == "NHS Ayrshire and Arran") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "NHS Ayrshire and Arran") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "NHS Ayrshire and Arran") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))

beds_oos %>%
  filter(hb_name == "NHS Ayrshire and Arran") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))
```


### Sanity check No. 2.
Check to see if I pick the aggregated results of value SB0801 in the location column
```{r}
beds_oos %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(hb_name == "The Golden Jubilee National Hospital" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

```{r}
beds_oos %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```


### Selected Geography: Special Health Board
```{r}
beds_oos %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))

beds_oos %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = hb_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))
```

## Calculate mean values


```{r}
kpi_mean_stats <- beds_oos %>% 
  group_by(hb) %>%
  # group_by(hb, hb_name) %>%
  filter(location == hb) %>% 
  summarise(
    # location_count = n(),
    mean_asb = mean(all_staffed_beds),
    mean_tob = mean(total_occupied_beds),
    mean_aasb = mean(average_available_staffed_beds), 
    mean_aob = mean(average_occupied_beds), 
    mean_po = mean(percentage_occupancy)
  )
```



# Join bed tibble with KPI stats tibble
```{r}
beds <- beds %>%
  left_join(kpi_mean_stats, "hb")

glimpse(beds)
```

# Save cleaned data

Write updated bed data to a CSV file.
```{r}
 write_csv(beds, here("beds_clean.csv"))
```



# KPI plots with stats (Scotland: 2020Q1 - 2021Q3)


## Selected Geography: Scotland
```{r}
beds %>%
  filter(hb_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = all_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = all_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = mean_tob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

beds %>%
  filter(hb_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = mean_tob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = mean_aasb), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = mean_aob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "Scotland") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_point(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_line(aes(y = mean_po), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


```


## Selected Geography: NHS Board
```{r}
beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = all_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = all_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = mean_tob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))

beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = mean_tob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = mean_aasb), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = mean_aob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_point(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_line(aes(y = mean_po), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))
```


## Selected Geography: Special Health Board
```{r}
beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = all_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = all_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = mean_tob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = total_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = mean_tob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_point(aes(y = average_available_staffed_beds, colour = hb_name)) +
  geom_line(aes(y = mean_aasb), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))



beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_point(aes(y = average_occupied_beds, colour = hb_name)) +
  geom_line(aes(y = mean_aob), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))


beds %>%
  filter(hb_name == "The Golden Jubilee National Hospital") %>% 
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, group = hb_name) +
  geom_line(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_point(aes(y = percentage_occupancy, colour = hb_name)) +
  geom_line(aes(y = mean_po), colour = "red") +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA))
```



# Split master beds tibble for shiny app

```{r}
# Load data
beds <- read_csv("beds_clean.csv") 

glimpse(beds)
```

## KPI tibble
```{r}
kpi_beds <- beds %>% 
  select(
    quarter, hb, hb_name, location,
    all_staffed_beds, total_occupied_beds, average_available_staffed_beds,
    average_occupied_beds, percentage_occupancy) %>% 
  rename(
    Quarter = quarter, 
    HB = hb, 
    HBName = hb_name, 
    Location = location,
    `All Staffed Beds` = all_staffed_beds, 
    `Total Occupied Beds` = total_occupied_beds, 
    `Average Available Staffed Beds` = average_available_staffed_beds,
    `Average Occupied Beds` = average_occupied_beds,
    `Percentage Occupancy` = percentage_occupancy
  )

```


## KPI differences tibble
```{r}
kpi_diff_beds <- beds %>% 
  mutate(
    all_staffed_beds0 = 100 * (all_staffed_beds - mean_asb) / mean_asb,
    total_occupied_beds0 = 100 * (all_staffed_beds - mean_tob) / mean_tob, 
    average_available_staffed_beds0 = 100 * (average_available_staffed_beds - mean_aasb) / mean_aasb,
    average_occupied_beds0 = 100 * (average_occupied_beds - mean_aob) / mean_aob, 
    percentage_occupancy0 = 100 * (percentage_occupancy - mean_po) / mean_po
  ) %>% 
  select(
    quarter, hb, hb_name, location,
    all_staffed_beds0, total_occupied_beds0, average_available_staffed_beds0,
    average_occupied_beds0, percentage_occupancy0)
```


```{r}
kpi_diff_beds <- kpi_diff_beds %>% 
  rename(
    Quarter = quarter, 
    HB = hb, 
    HBName = hb_name, 
    Location = location,
    `All Staffed Beds` = all_staffed_beds0, 
    `Total Occupied Beds` = total_occupied_beds0, 
    `Average Available Staffed Beds` = average_available_staffed_beds0,
    `Average Occupied Beds` = average_occupied_beds0,
    `Percentage Occupancy` = percentage_occupancy0
  )
```




# Save KPI data

Write updated bed KPI data to a CSV file.
```{r}
 write_csv(kpi_beds, here("kpi_beds_clean.csv"))

 write_csv(kpi_diff_beds, here("kpi_diff_beds_clean.csv"))
```




# Doodles

<!-- ```{r} -->
<!-- country_name_beds <- unique(beds$country_name) %>%  -->
<!--   discard(is.na) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- hb_name_beds <- unique(beds$hb_name) %>% -->
<!--   discard(is.na) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- shb_name_beds <- unique(beds$shb_name) %>%  -->
<!--   discard(is.na) -->

<!-- ``` -->
